---
title: "Portfolio Optimization"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(gganimate)
library(gifski)
```

```{r}
f = NULL
files = c("TSLA.csv","AMZN.csv","AAPL.csv","SONO.csv","SBSW.csv","GFI.csv")
for (i in 1:length(files)) {
  csv = read.csv(files[i])
  csv = csv[,c("Date","Close")]
  names(csv) = c("Date",files[i])
  if (i == 1) f = csv
  else f = merge(f,csv)
}
```

Calculate Asset Returns

```{r}
for (i in 2:ncol(f)) {
  # Price time series of the i-th asset
  prices = f[,i] 
  
  # Price lagged by 1
  prices_prev = c(NA,prices[1:(length(prices)-1)]) 
  
  # Returns time series
  returns = (prices-prices_prev)/prices_prev 
  
  # Replace the i-th column with returns
  f[,i] = returns 
}
# Remove the first row with NAs and the Date column
asset_returns = f[2:nrow(f),2:ncol(f)]
```

## Portfolio Returns

Portfolio historical returns are a function of the weights. Remember that historical data is always the same and what we want to do is to find an optimal portfolio **changing the weights values**.

```{r}
portfolio_returns = function(x) {
  port.returns = 0
  
  #Multiplication of the i-th asset by the i-th weight in "x"
  for (i in 1:length(x)) {
    port.returns = port.returns + asset_returns[,i] * x[i]
    
  }
  return (port.returns)
}
```

## Objective function with Penalty

First of all, we have to calculate the Sharpe ratio on the historical weighted portfolio returns.

```{r}
sharpe = function(x) {
  port.returns = portfolio_returns(x)
  
  return (mean(port.returns)/sqrt(var(port.returns)))
  
}
```

Now we have to write the code for the **penalty function**. We can write a *constraint* function that implements all the constraints.

```{r}
constraint = function(x) {
 boundary_constr = (sum(x)-1)**2   # "sum x = 1" constraint
 
 for (i in 1:length(x)) {
  boundary_constr = boundary_constr + 
     max(c(0,x[i]-1))**2 +  # "x <= 1" constraint
     max(c(0,-x[i]))**2     # "x >= 0" constraint
 }
 
 return (boundary_constr)
}
```

```{r}
obj = function(x) {
  #we want to maximize the Sharpe Ratio
  return (-sharpe(x) + 100*constraint(x))
}
```

## Optimization via Genetic Algorithm

The *ga* function in R is a **simple but effective** genetic algorithm implementation for solving maximization problems.

For this article, I configured the genetic optimization in order to make it perform 50.000 iterations, stopping only if the maximum fitness doesn't change for 50 consecutive iterations. This is our **stopping** criterion for the genetic algorithm.

```{r}

library("GA")
ga_res = ga(
      # Tell the genetic algorithm that the 
      # weights are real variables
      type="real-valued", 
      
      # "ga" function performs maximization, so we must
      # multiply the objective function by -1
      function(x){-obj(x)}, 
      
      # x_i >= 0
      lower = rep(0,ncol(asset_returns)), 
      
      # x_i <= 1
      upper = rep(1,ncol(asset_returns)), 
      
      # Maximum number of iterations 
      maxiter = 50000, 
      
      # If the maximum fitness remains the same for 50
      # consecutive transactions, stop the algorithm
      run=50, 
      
      keepBest = TRUE,
      
      # Exploit multi-core properties of your CPU
      parallel=TRUE,
      
      # We want to see the partial results of the process
      # while it performs
      monitor=TRUE,
      
      # Seed useful for replicating the results
      seed=1
)
```

Let's store the resulting weights in a vector.

```{r}
# Store the resulting weights in a vector
sol = as.vector(summary(ga_res)$solution)
optimal_returns = portfolio_returns(sol)
```

```{r}

df1 <- data.frame(cumsum(optimal_returns))
df <- cbind(df1,cumsum(asset_returns))
colnames(df)[1]  <- "Weighted Portfolio"
```

```{r}

df$idu <- row.names(df)
df$idu <- as.numeric(row.names(df))

df %>% 
  gather(var, value, -idu) %>% 
  ggplot(aes(idu, value, group = var, color = var)) + 
  geom_line() +
  labs(color='Stock', x = "Index", y ="Return") 

```

## Part 2

we want to take n number of weighter portfolios and the last one, make GGplots with each one of them and export images to a folder.

Then create a function to import these images and play as a movie

```{r}
All_solutions <- ga_res@bestSol

```

```{r}
Test_Sol <- All_solutions[1:50] 
```

Determine a list of all Optimal returns

```{r}
sol_M = as.vector(Test_Sol)

optimal_returns_M <-0
for (i in 1:length(sol_M)) {
  optimal_returns_M <- cbind(optimal_returns_M, portfolio_returns(sol_M[[i]]))
}
  
optimal_returns_M <- optimal_returns_M[,-1] 
```

Create data frame to plot

```{r}
Cum_Asset_returns <- cumsum(asset_returns)
Cum_Asset_returns$index <- row.names(Cum_Asset_returns)
Cum_Asset_returns$index <- as.numeric(row.names(Cum_Asset_returns))


Cum_opt_returns_M <- 0 

for (i in 1:ncol(optimal_returns_M)) {
  Cum_opt_returns_M <- cbind(Cum_opt_returns_M, cumsum(optimal_returns_M[,i]))
  
}


Cum_opt_returns_M <- Cum_opt_returns_M[,-1] 
#Cum_opt_returns_M <- cbind(Cum_opt_returns_M,Cum_Asset_returns$index)

```

Make plots

```{r}
library("reshape2")


for (i in 1:(ncol(Cum_opt_returns_M)-1)) {
  df4 <- cbind(Cum_opt_returns_M[i],Cum_Asset_returns)
  df4 <- melt(df4, id.vars="index", value.name="value", variable.name="stock")
  
  myPlot <- ggplot(df4, aes(x=index,y= value, color = stock)) +
    geom_line() +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_along}', x = 'GDP per capita', y = 'life expectancy') +
  #transition_time(index) +
  transition_reveal(df4$index)+
  ease_aes('linear')

  
}


```

```{r}
animate(myPlot, duration = 5, fps = 20, width = 800, height = 600, renderer = gifski_renderer(), rewind = FALSE)
#anim_save("output.gif")
```

```{r}
library("reshape2")



for (i in 1:(ncol(Cum_opt_returns_M)-1)) {
  df4 <- cbind(Cum_opt_returns_M[,i],Cum_Asset_returns)
  df4 <- melt(df4, id.vars="index", value.name="value", variable.name="stock")
  
  myPlot[[i]] <- ggplot(df4, aes(x=index,y= value, color = stock)) +
    geom_line() +
  labs(title = 'T', x = 'GDP per capita', y = 'life expectancy')

}

```

```{r}
library(animation)

# explicit
animation::saveGIF(
  expr = {
    for (i in 1:length(myPlot)) {
      plot(myPlot[[i]])
    }
    

  },
  movie.name = "explicit_my3.gif",fps =2
)

```
